Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
rename annales folder, bigoplus
fix some equations in surfaces, line breaks
some hirz generators rewording

Another closely related topic is the Hassett-Keel program 
initiated in \todo{was it actually initiated here? check vzb.} \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \oplus_{d \geq 0}H^0 \left( \overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} + \alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces. Fractional divisors on surfaces also naturally appear when considering the canonical ring of surfaces of Kodaira dimension 1. A surface has Kodaira dimension 1 if and only if it is an elliptic surface, and their canonical ring are most naturally described in the setting of fractional divisors. \todo{Add references}

@@

\begin{thm}
\label{lem:hirz-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\hirz_k$.  Let $C_i = V(f_i)$ where $f_i \in \mathscr{P}(a_i, b_i)$; further suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.
Then $R_D$ is generated in degree less than 
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{i\in \Tp(D) \atop
	j \in \Tm(D)} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}
\]
with relations generated in degree less than 
\[
	\max(2 \sigma, \sigma + \tau)
\]
where
\[
	\tau = \sigma
	+ \max( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{i\in S_+ \atop j\in S_-}(\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))).
\]
\end{thm}

@@

To do this, we extend the method of O'Dorney \cite[Theorem 8]{dorney:canonical}. 
We first
view the cone inside of $\mathbb{R}^{n+1}$ given by elements of the
form $(d, c_1, \ldots, c_n)$ satisfying $c_i \ge -d \alpha_i$ for
all $i$ and $\sum_{i = 1}^n c_i (\bida_i + \bidb_i) =0$. By O'Dorney, this
has extremal rays given by

@@

Let $\Delta^{n-1}$ be the simplex with vertices $e_1, \ldots e_n$.
We can intersect $\Delta^{n-1}$ with the hyperplane $H$ given by 
$\sum_{i=1}^n \bida_i X_i = 0$
 to get a subspace $S \subseteq \Delta^{n-1}$. The extremal rays of $S$ must lie on the edges of $\Delta$ \todo{Peter: Do we need to cite/argue this, or is it clear?} and are thus given
by the set of $e_i$ contained in $H$ together with the points $e_{i, j}$ given by intersections of $H$ with edges of the simplex
connecting $e_i, e_j \not \in H$.
\todo{more detail above}

@@

\begin{lem}
$\ker(\psi)$ is generated in less than
\[
	\tau = \sigma
	+ \max( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{i\in S_+ \atop j\in S_-}(\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))).
\]
\end{lem}

>>>>

Another closely related topic is the Hassett-Keel program 
initiated in \todo{was it actually initiated here? check vzb.} \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0}H^0 \left( \overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} + \alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces. Fractional divisors on surfaces also naturally appear when considering the canonical ring of surfaces of Kodaira dimension 1. A surface has Kodaira dimension 1 if and only if it is an elliptic surface, and their canonical ring are most naturally described in the setting of fractional divisors. \todo{Add references}

@@

\begin{thm}
\label{lem:hirz-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\hirz_k$.  
Let $C_i = V(f_i)$ where $f_i \in \mathscr{P}(a_i, b_i)$; further 
suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.
Then $R_D$ is generated in degree less than 
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{i
	\in \Tp(D) \atop
	j \in \Tm(D)} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}
\]

\noindent
with relations generated in degree less than 
\[
	\max(2 \sigma, \sigma + \tau)
\]

\noindent
where
\[
	\tau = \sigma
	+ \max \left( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{i \in
	S_+ \atop j\in S_-}(\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))
	\right).
\]
\end{thm}

@@

To do this, we extend the method of O'Dorney \cite[Theorem 8]{dorney:canonical}. 
We first view the cone inside of $\mathbb{R}^{n+1}$ given by elements of the
form $(d, c_1, \ldots, c_n)$ satisfying $c_i \ge -d \alpha_i$ for
all $i$ and $\sum_{i = 1}^n c_i (\bida_i + \bidb_i) = 0$. By O'Dorney, this
has extremal rays given by

@@

Let $\Delta^{n-1}$ be the simplex with vertices $e_1, \ldots e_n$.
We can intersect $\Delta^{n-1}$ with the hyperplane $H$ given by
$\sum_{i=1}^n \bida_i X_i = 0$  to get a subspace $S \subseteq \Delta
^{n-1}$. The extremal rays of $S$ must lie on the edges of $\Delta
^{n - 1}$ \todo{Peter: Do we need to cite/argue this, or is it clear?}.
Thus, the extremal rays of $S$ are given by the set of $e_i$
contained in $H$ and by the points $e_{i, j}$ that are the
intersections of $H$ with edges of the simplex connecting the
distinct $e_i, e_j$ not contained in $\in H$.
\todo{more detail above}

@@

\begin{lem}
$\ker(\psi)$ is generated in less than
\[
	\tau = \sigma
	+ \max \left( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{i \in
	S_+ \atop j\in S_-}(\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))
	\right).
\]
\end{lem}
